Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept

### 2021-11-18 paper/article.tex
Paper progress.

\title[Modeling a Neural Co-Processor]
{Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept}

@@

An example of such a simulation approach is Bernal et al. \cite{bernal.sim}. In this work, the
authors use a simulated spiking neural network to train a stimulation agent, much
like the work we present here. Their stimulation agent sought to restore the network's
control of a simulated arm, to reach a target, after a simulated lesion was applied.
As they point out, we have a limited ability to probe a neural circuit \textit{in vivo}
in order to perform learning. As a result, we first need to design our approach through
the use of an admissible simulation. The authors in this case simulated lesions by
effectively removing parts of their simulated network, or by cutting connections between
parts of the network. Likewise here, we use an established design for an artificial
neural circuit which performs a grasping task, whose architecture and training methods
were designed specifically to result in naturalistic dynamics.

@@

TODO: summary of conclusions that will go here:
\begin{itemize}
	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
		\begin{itemize}
			\item 90\% on connection lesion, both with and without co-adaptation
			\item M1 50\%: 81\% recovery on first attempt with co-adaptation
			\item AIP 50\%: 35\% recovery w/co-adaptation. Presumably, this is because we've lost encoding functionality, and co-adaptation is very slow. If we ran the simulation for 1 month, who knows?
			\item Waiting on all other results.
		\end{itemize}
	\item Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
	\item Fine tuning takes a far longer time
	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
	\item Can co-adapt with the brain, as it changes
	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
	\item Show pre-recovery results; co-processor will improve beyond what recovery can achieve. (will run results this week)
\end{itemize}

@@

We present here a first-of-its-kind demonstration of a novel design for neural co-processors,
allowing for an AI agent to learn stimulation policies that improve a user's task performance.
The central notion of the design is to leverage a stimulation model, whose purpose is to provide
training feedback to an AI agent responsible for generating the stimulation. We present here
a first demonstration of the learning algorithms and key properties

We proposed X
It did Y
It was found that Z
It has potential to A

@@

One clear advantage of our co-processor approach is that, as in Bernal \cite{bernal.sim},
we can train our stimulation agent to a great degree

>>>>

\title[Modeling a Neural Co-Processor]
{Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept}

@@

An example of such a simulation approach is Dura-Bernal et al. \cite{bernal.sim}. In this work, the
authors use a simulated spiking neural network to train a stimulation agent, much
like the work we present here. Their stimulation agent sought to restore the network's
control of a simulated arm, to reach a target, after a simulated lesion was applied.
As they point out, we have a limited ability to probe a neural circuit \textit{in vivo}
in order to perform learning. As a result, we first need to design our approach through
the use of an admissible simulation. The authors in this case simulated lesions by
effectively removing parts of their simulated network, or by cutting connections between
parts of the network. Likewise here, we use an established design for an artificial
neural circuit which performs a grasping task, whose architecture and training methods
were designed specifically to result in naturalistic dynamics.

@@

TODO: summary of conclusions that will go here:
\begin{itemize}
	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
		\begin{itemize}
			\item 90\% on connection lesion, both with and without co-adaptation
			\item M1 50\%: 81\% recovery on first attempt with co-adaptation. This is likely to go higher, since I found a bug.
			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running. The loss was very low to start with, and recovery depends heavily on co-adaptation, so training is slow. Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
			\item Waiting on all other results.
		\end{itemize}
	\item Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
	\item Fine tuning takes a far longer time
	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
	\item Can co-adapt with the brain, as it changes
	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
	\item Show pre-recovery results; co-processor will improve beyond what recovery can achieve. (will run results this week)
\end{itemize}

@@

We present here a first-of-its-kind demonstration of a novel design for neural co-processors,
allowing for an AI agent to learn neural stimulation policies that improve a user's performance
of an external task. The design revolves around the use of a stimulation model for training the
agent, providing a proxy to the true function mapping stimulation and neural activity
to task performance. We base our demonstration on a simulation of a neural circuit engaged in
an external grasping task.

Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
50-90\% across our various experiments. To work this well, the co-processor needed to adapt to
the long-running dynamics of the stimulated network, as well as the long-running effects of
stimulation. In some experiments we required it to additionally adapt to non-stationarity in
the neural circuit, which was actively changing at the same time the co-processor was learning.
It achieved that as well, co-adapting with the simulated brain to improve external task
performance.  TODO: pre-recovery result.

Clearly, we need to examine carefully what can and cannot be inferred from simulation studies
such as these. As in Dura-Bernal \cite{bernal.sim}, we note that our simulation differs
drastically from an organic brain, both in scale and structure. Today, predicting long-running
neural responses to stimulation remains a difficult problem, implying targeted neural control
is also difficult. Even moreso, identifying the neural correlates of complex task success
remains largely out of reach \cite{khanna.openloop}. What, then, does our simulation provide us?

Abstractly, the co-processor lends itself to any closed-loop neural stimulation problem
where a relevant stimulation model can be identified. There are many lower dimensional problems
today which we address with far simpler stimulation regimes than would be necessary for
e.g. remediating the symptoms of a stroke. For example, deep brain stimulators for
treating the symptoms of Parkinson's Disease can now be controlled in a closed loop
fashion to also minimize stimulation side-effects and maximize implant battery life.
TODO: citation. Our simulation clearly does not imply that our co-processor design can
translate immediately to stroke applications. However, it does demonstrate that our design
can adapt to long-running effects of stimulation, as well as to the non-stationarity of
the neural circuit to which it is applying the stimulation. These properties together
suggest it could be useful in lower dimensional problems where a stimulation model can
be estimated, and stimulation parameters can be tied to the target objective. TODO: more examples?

On the whole, the co-processor quickly improved task performance, but required orders of magnitude
more training examples to achieve its highest levels of performance. This training inefficiency remains
a problem...

@@

One clear advantage of our co-processor approach is that, as in Dura-Bernal \cite{bernal.sim},
we can train our stimulation agent to a great degree
